---
title: "Project 3 - Data Science Skills"
author: "David Apolinar, Anthony Muñoz, Christina Valore, Omar Pineda."
date: "3/24/2019"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    source_code: embed
---

Row {.tabset .tabset-fade}
-------------------------------------

### Overview

Data was brough in two ways: via scraping and manual collection. The manual collection was taken from LinkedIn profiles, where we collected 40 data scientists information including: education, school, location,and skills. For the scraping, we used the kdnuggets website to look at tools currently used by data scientist's and trends from 2017 - 2018. 

The LinkedIn data was then added to a relational database and tidied inside of R. The scraped data was brought directly into R. After the data was in their respective dataframes, we then graphed, created wordclouds and used a map to analyze the data. For the graphing, we made use of the ggplot library and for the map we used ggmap package.

### Data Load

Several of the top data scientist's information was stored in in a CSV file and was tidy up using dplyr libraries.

```{r data load}
library(tidyr)
library(wordcloud)
library(tm)
library(SnowballC)
library(RColorBrewer)
library(magrittr)
library(DBI)
library(dplyr)
library(rvest)
library(stringr)
library(ggplot2)
library(RColorBrewer)
#  Data Files
ds <- read.csv("https://raw.githubusercontent.com/omarp120/DATA607Week8/master/DataScientists.csv", header=TRUE, stringsAsFactors = FALSE)
tidyDS <- gather(ds, "Number", "Skill", Skill1:Skill50) #makes data tall
finalDS <- tidyDS[tidyDS$Skill != "",] #removes rows with empty skill values
counts <- as.data.frame(table(finalDS$Skill)) #creates a data frame with skill frequencies

```

### SQL Creation and Insertion

DBI was used to create tables on the fly with an Azure SQL DB for MySQL. The databases were created from R and all insert statements were completed from R. The following tables were created:

Skills Table
Person Table

The tables were created with a Many to Many relationship.

```{r SQL dataload}
# Create Skill Table
skilltable <- unique(finalDS$Skill)
skilltable <- as.data.frame(skilltable, stringsAsFactors = FALSE)

skillids <- 1:nrow(skilltable)

skilltable <- cbind.data.frame(skilltable,skillids)
names(skilltable) <- c("SkillName", "SkillID")

# Run SQL statements to create tables 

con <- dbConnect(RMariaDB::MariaDB(), user='x-admin@cunyspsmysql.mysql.database.azure.com', password="7dWa0XUVHtlCJMS", dbname='datascientists' ,host='cunyspsmysql.mysql.database.azure.com')

rs<- dbSendStatement(con, "drop table if exists person_skills;")
dbClearResult(rs)
rs<-dbSendStatement(con, "drop table if exists person;")
dbClearResult(rs)
rs<-dbSendStatement(con, "drop table if exists skills;")
dbClearResult(rs)

rs <- dbSendStatement(con, "CREATE TABLE person (
                personid int NOT NULL auto_increment primary key,
                title nchar(50),
                name nchar(50) NOT NULL,
                education nchar(50),
                degree nchar(50),
                location nchar(50),
                company nchar(50));")
dbClearResult(rs)
rs<- dbSendStatement(con, "CREATE TABLE skills (
                skillid int NOT NULL auto_increment primary key,
                skillname nchar(50) NOT NULL);")
dbClearResult(rs)
rs<- dbSendStatement(con, "CREATE TABLE person_skills (
personid int NOT NULL references person(personid),
                skillid int NOT NULL references skills(skillid),
                CONSTRAINT person_skill primary key(personid, skillid));")
dbClearResult(rs)
dbDisconnect(con)
# Create SQL Connection

con <- dbConnect(RMariaDB::MariaDB(), user='x-admin@cunyspsmysql.mysql.database.azure.com', password="7dWa0XUVHtlCJMS", dbname='datascientists' ,host='cunyspsmysql.mysql.database.azure.com')

#mysql_datascientists <- dbGetQuery(con, 'select * from skills')
for(i in 1:nrow(skilltable))
{
#  print(paste0("Inserting Skill: ", skilltable[i,]$SkillName, ", SkillID: ", skilltable[i,]$SkillID) )
  sql <- sprintf("insert into skills
                  (skillname, skillid)
               values ('%s', %d);",
               skilltable[i,]$SkillName, skilltable[i,]$SkillID)
  rs <- dbSendQuery(con, sql)
  dbClearResult(rs)
}

mysql_dataskills <- dbGetQuery(con, 'select * from skills
                               limit 10')
mysql_dataskills
dbDisconnect(con)
```

Below is a list of the people being pulled from a database.

mysql_datascientists <- dbGetQuery(con, 'select name,education,degree,title,company,location from person
                                   limit 10;')
We limit the results to only 10 rows.

```{r insert peope}
# Get Unique People to Insert
con <- dbConnect(RMariaDB::MariaDB(), user='x-admin@cunyspsmysql.mysql.database.azure.com', password="7dWa0XUVHtlCJMS", dbname='datascientists' ,host='cunyspsmysql.mysql.database.azure.com')

people_table <- finalDS %>% select(ID,Person, Title, School, HighestLevel, Location, Company) %>% unique()

for(i in 1:nrow(people_table))
{
#  print(paste0("Inserting Person: ", 
#               people_table[i,]$Person, ", Title: ", 
#               people_table[i,]$Title, "School: ", 
#               people_table[i,]$School, ", Degree: ", 
#               people_table[i,]$HighestLevel, ", Location: ", 
#               people_table[i,]$Location, ", Company: ",
#               people_table[i,]$Company))
  sql <- sprintf("insert into person
                 (name, title, education, degree, location, company)
                 values ('%s', '%s', '%s','%s', '%s', '%s');",
                 people_table[i,]$Person, 
                 people_table[i,]$Title, 
                 people_table[i,]$School,
                 people_table[i,]$HighestLevel, 
                 people_table[i,]$Location,
                 people_table[i,]$Company)
  rs <- dbSendQuery(con, sql)
  dbClearResult(rs)
}

mysql_datascientists <- dbGetQuery(con, 'select name,education,degree,title,company,location from person
                                   limit 10;')
mysql_datascientists
dbDisconnect(con)



# Create Many to Many Relationship
linkdb<- tidyDS %>% select(ID, Skill)
returnIndex <- function(n)
{
  for(i in 1:nrow(n))
  {
    
    return (skilltable$SkillID[skilltable$SkillName == n[i,]$Skill])
  }

}
# Remove duplicate rows
person_skill <- finalDS %>% select(ID, Person, Skill) %>% distinct()

#returnIndex(linkdb[478,])

# Create Link Table
con <- dbConnect(RMariaDB::MariaDB(), user='x-admin@cunyspsmysql.mysql.database.azure.com', password="7dWa0XUVHtlCJMS", dbname='datascientists' ,host='cunyspsmysql.mysql.database.azure.com')


for(i in 1:nrow(person_skill))
{
  if(length(returnIndex(person_skill[i,])) != 0)
  {
#    print(paste0("Inserting (PersonID: ", person_skill[i,]$ID, " SkillID: ", returnIndex(person_skill[i,]),")") )
    
    sql <- sprintf("insert into person_skills
                 (personid, skillid)
                   values (%d, %d);",
                   person_skill[i,]$ID, returnIndex(person_skill[i,]))
    rs <- dbSendQuery(con, sql)
    dbClearResult(rs)
  }else
  {
    print("Empty Skill Value, skipping link")
  }
}

dbDisconnect(con)
```

### Web Scraping Information


On this section we did web scraping on the KDnuggets website where we gather some information about a Software poll about what’s tools in percent the software developer and data Scientist have share and use more. we can look at the major analytics tools, language programing and others tools and also the increasing and decreasing of those tools between years 2017-2018.

```{r}

data <- read_html("https://www.kdnuggets.com/2018/05/poll-tools-analytics-data-science-machine-learning-results.html")
table1 <- data %>% html_nodes("table") %>% .[1] %>% html_table(fill = T) %>% as.data.frame() %>% lapply( gsub, pattern='%', replacement='' ) %>% as.data.frame()  %>% droplevels()
colnames(table1) <- c("Software", "2018 % share","% change 2018 vs 2017")
#as.numeric(as.character(table1$`2018 % share`))
table2 <- data %>% html_nodes("table") %>% .[2] %>% html_table(fill = T) %>% as.data.frame() %>% lapply( gsub, pattern='%', replacement='') %>% as.data.frame()
colnames(table2) <- c("Tool", "% change","2018 % share","2017 % share")
table3 <- data %>% html_nodes("table") %>% .[3] %>% html_table(fill = T) %>% as.data.frame() %>% lapply( gsub, pattern='%', replacement='') %>% as.data.frame()
colnames(table3) <- c("Tool", "% change","2018 % share","2017 % share")
table4 <- data %>% html_nodes("table") %>% .[4] %>% html_table(fill = T) %>% as.data.frame() %>% lapply( gsub, pattern='%', replacement='') %>% as.data.frame()
colnames(table4) <- c("Tool", "% change","2018 % share","2017 % share")
```


### Web Scraping analysis

```{r web scrape analysis}

#Table 1: Top software
ggplot(table1,aes(x=Software, y=`% change 2018 vs 2017`,fill= Software)) + geom_bar(stat="identity") + scale_fill_brewer(palette="Set3") + xlab("Software") + ylab("Percent Change") +  ggtitle("Top Analytics Software in 2018") + theme(legend.position = "none",  
            axis.text.x = element_text(angle = 25, hjust = 1))

#Table 2: Top software with increase in usage
ggplot(table2,aes(x=Tool, y=`% change`, fill = Tool)) + geom_bar(stat="identity")+ scale_fill_brewer(palette="Paired") + xlab("Software") +  ylab("Percent Change") +  ggtitle("Major Analytics Tools with largest increase from 2017 - 2018")  + theme(legend.position = "none", axis.text.x = element_text(angle = 25, hjust = 1))

#Table 3: Top software with decline in usage

#n is large so we must increase the palatte size
colorCount = length(unique(table3$Tool))
getPalette = colorRampPalette(brewer.pal(9, "Set1"))

ggplot(table3,aes(x=Tool, y=`% change`, fill = Tool)) + geom_bar(stat="identity", fill=getPalette(colorCount)) + xlab("Software") +  ylab("Percent Change") +  ggtitle("Major Analytics Tools with largest decline from 2017 - 2018")  + theme(legend.position = "none", axis.text.x = element_text(angle = 25, hjust = 1))

#Table 4: Big data tools decline
ggplot(table4,aes(x=Tool, y=`% change`, fill = Tool)) + geom_bar(stat="identity")+ scale_fill_brewer(palette="Set2") + xlab("Software") +  ylab("Percent Change") +  ggtitle("Decline in Big Data Tools Usage from 2017 - 2018")  + theme(legend.position = "none", axis.text.x = element_text(angle = 25, hjust = 1))

```

### Word Cloud

Using a word cloud graph, we can visually see that the most common skills amoung the top data scientists are Machine Learning, Data Mining, Python, Data Analysis, and Algoritms.

```{r data vis}
colnames(counts) <- c("Skill", "Freq")
wordcloud(counts$Skill, counts$Freq, random.order = FALSE, scale = c(2, 0.10), colors=brewer.pal(8, "Dark2"))
```

### Data Scientists Common Locations

```{r goecode}

#code adapted from http://www.storybench.org/geocode-csv-addresses-r/
#library(ggmap)
#register_google(key = "xxx") #removed personal API key
# Initialize the data frame
#getOption("ggmap")
# Loop through the addresses to get the latitude and longitude of each address and add it to the
# origAddress data frame in new columns lat and lon
#for(i in 1:nrow(ds))
#{
  # Print("Working...")
#  result <- geocode(ds$Location[i], output = "latlon", source = "google")
#  ds$lon[i] <- as.numeric(result[1])
 # ds$lat[i] <- as.numeric(result[2])
#}
# Write a CSV file containing origAddress to the working directory
#write.csv(ds, "geocoded.csv", row.names=FALSE)
```

```{r map}
library(leaflet)
cities <- read.csv("https://raw.githubusercontent.com/omarp120/DATA607Week8/master/geocoded.csv")
cities  %>% 
  leaflet() %>% 
  addTiles() %>% 
  addMarkers(clusterOption=markerClusterOptions())
```

### Top Skills Analysis

```{r skill graph}

# grouped by skill and summarized into new data frame
skill_df<-finalDS %>%
  group_by(Skill) %>%
  summarize(n())

skill_df_top<-arrange(skill_df,desc(`n()`))

top10<- head(skill_df_top,10)

ggplot(top10,aes(x=Skill, y=`n()`,fill= top10$Skill)) + geom_bar(stat="identity") + scale_fill_brewer(palette="Spectral") + xlab("Skill") + ylab("Count") +  ggtitle("Top 10 Data Scientist Skills ") + theme(legend.position = "none",  
            axis.text.x = element_text(angle = 25, hjust = 1))
```

### Conclusion

For the top data scientist skills, we can see machine learning followed by data mining and Python are at the top of the list. This could be due to the increased demand of predicitive analytics. In terms of datat science software, there is a decrease in the use of R, while there is a signifigant increase in RapidMiner from 2017 to 2018. Other tools with increased adoption from 2017 - 2018 are Keras and PyTorch, while Caffe and Machine Learning Server saw over a 50% decline in usage. 

### References

1-) Web Scrapping website,data retrieved 3/21/2019
https://www.kdnuggets.com/2018/05/poll-tools-analytics-data-science-machine-learning-results.html
