---
title: "Week7_WebTech_OP"
author: "Omar Pineda"
date: "3/11/2019"
output: html_document
---

In this assignment, we will practice loading unstructured data from different file types. First, we load in our books data from an HTML table:

```{r htmlLoad}
library(RCurl)
library(XML)
link <- "https://raw.githubusercontent.com/omarp120/DATA607Week7/master/books.html"
booksHTML <- readHTMLTable(getURL(link))[[1]]
booksHTML
```

Next, we load these same books from a JSON file:

```{r jsonLoad}
library(jsonlite)
link2 <- "https://raw.githubusercontent.com/omarp120/DATA607Week7/master/books.json"
booksJSON <- fromJSON(link2)[[1]]
booksJSON
```

Finally, we load these books from an XML file:

```{r xmlLoad}
library(XML)
library(plyr)
link3 <- "https://raw.githubusercontent.com/omarp120/DATA607Week7/master/books.xml"
books <- xmlParse(getURL(link3), useInternalNodes = TRUE, validate = FALSE)
booksXML <- ldply(xmlToList(books), data.frame)
booksXML <- booksXML[,2:5] #removes first column, which is an id
booksXML
```

The tables are slightly different, some containing additional columns than others. Also, the XML and HTML tables' elements have data type 'fctr' while the JSON table's elements are 'chr'.