---
title: "Week3_Strings_OP"
author: "Omar Pineda"
date: "2/13/2019"
output: html_document
---

3. Copy the introductory example. The vector name stores the extracted names.

```{r loading}
library(stringr)
raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
name
```
(a) Use the tools of this chapter to rearrange the vector so that all elements conform to
the standard first_name last_name.

First, I'll separate those names that are already in this format from those that need to be re-arranged.

```{r name}
rightOrder <- c(name[1], name[3], name[4], name[6])
wrongOrder <- c(name[2], name[5])
wrongOrder
```

Now, we change the order of the names that are not in the correct order.

```{r switch}
montgomery <- unlist(str_split(wrongOrder[1], ","))
montgomery
montgomery <- str_trim(montgomery)
montgomery
add1 <- str_c(montgomery[2], montgomery[1], sep=" ")
add1
homer <- unlist(str_split(wrongOrder[2],","))
homer <- str_trim(homer)
add2 <- str_c(homer[2], homer[1], sep=" ")
```

Finally, we add these two corrected names to the vector that contains names with the right order.

```{r finalNames}
rightOrder <- c(rightOrder, add1, add2)
rightOrder
```

(b) Construct a logical vector indicating whether a character has a title (i.e., Rev. and
Dr.).

```{r title}
hasTitle <- str_detect(rightOrder, "[aA-zZ]{2,}\\.")
hasTitle
```

(c) Construct a logical vector indicating whether a character has a second name.

```{r second}
hasSecondName <- str_detect(rightOrder, "[A-Z]{1}\\.")
hasSecondName
```

4. Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.

(a) [0-9]+\\$
Strings that have a series of at least 1 integer followed by a dollar sign conform to this expression. Example: abc12847$a

(b) \\b[a-z]{1,4}\\b
Strings that contain 1-4 consecutive lowercase letters conform to this expression. Example: abc

(c) .*?\\.txt$
Strings that end in .txt conform to this expression. Example: ok.txt

(d) \\d{2}/\\d{2}/\\d{4}
Strings that have two digits, a forward slash, two more digits, another forward slash, and lastly 4 digits conform to this expression. These are dates. Example: 02/13/2019

(e) <(.+?)>.+?</\\1>
XML tags conform to this expression. This is a tag followed by text and ending in the same tag that it began with, except with a forward slash. Example: < address >This is correct syntax< / address > 